Supermartingales and Termination Analysis of Statistical PCF

### 2020-12-28 LICS21/SPCF.tex
(Global change: probability space (S, \calF, \mu) -> (\entrosp, \calF, \mu_\entrosp)

The language SPCF is a simply-typed lambda calculus with sampling of real numbers from $[0,1]$ and unbounded scoring, following \cite{MakOPW21}. 
Types and terms are defined as follows, where $r$ is a real number, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function, and $\Gamma$ is an environment:
\begin{align*}
  \text{types } A, B &::= \textsf{R}  \mid  A \to B \\
  \text{values } V &::= \lambda x.M  \mid  \underline{r} \\
  \text{terms } M, N &::= V  \mid  x  \mid  M_1 \, M_2  \mid  \underline{f}(M_1,\dots ,M_n)  \mid  \tY M \\
   & \qquad \mid  \tif{M < 0}{N_1}{N_2}  \mid  \tsample  \mid  \tscore(M)
\end{align*}
The typing rules are presented in \Cref{fig:typing rules}.
Terms are identified up to $\alpha$-equivalence, as usual. 
The set of all terms is denoted $\Lambda$, and the set of closed terms is denoted $\Lambda^0$.

@@

\begin{align*}
  E[(\lambda x.M) N] & \to E[M[N/x]] \\
  E[\underline f (\underline r_1, \dots , \underline r_n)] & \to E[\underline{f(r_1,\dots,r_n)}] \\
  E[\tY \lambda x. M] & \to E[\lambda z. M[(\tY \lambda x. M)/x] z] \\
  & \qquad \qquad \text{ where $z$ is not free in $s$}\\
  E[\tif{\underline r < 0}{M_1}{M_2}] & \to E[M_1] \text{ where }r < 0 \\
  E[\tif{\underline r < 0}{M_1}{M_2}] & \to E[M_2] \text{ where }r \geq 0 \\
  E[\tsample] & \to E[\underline r] \text{ where } r \in [0,1] \\
  E[\tscore(\underline r)] & \to E[\underline r].
\end{align*}
We write $\to^\ast$ for the reflexive, transitive closure of $\to$.
\lo{The $\tscore$ rule should have a side condition: if $r > 0$.}

@@

\subsection{Sampling semantics}
\label{sec:sampling semantics}
This version of the reduction relation allows $\tsample$ to reduce to any number in $[0,1]$. To more precisely specify the probabilities, an additional argument is needed to determine the outcome of random samples. Let $ I = [0,1] \subset \mathbb{R} $, and let $S = I^{\mathbb{N}}$, with the \changed[lo]{Borel $\sigma$-algebra $\calF$} and the probability measure, denoted $\mu$, given by the limit of $1 \gets I \gets I^2 \gets \cdots$, where the maps are the projections that ignore the last element. Equivalently, a basis of measurable sets is $\prod_{i=0}^\infty X_i$ where $X_i$ are all Borel and all but finitely many are $I$, and $\mu (\prod_{i=0}^\infty X_i) = \prod_{i=0}^\infty \Leb(X_i)$.
The maps $\pi_h:S \to I, \; \pi_t:S \to S$ popping the first element are then measurable.
\changed[lo]{Following \cite{DBLP:conf/esop/CulpepperC17}, we call the probability space $(S, \calF, \mu)$ the \emph{entropy space}.}

@@

The one-step reduction is given by the function $\red : \Lambda^0 \times S \to \Lambda^0 \times S$ where
\begin{equation*}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N\\
        & \qquad \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ is a value}
    \end{array} \right .
\end{equation*}
\lo{Both $\to$ and $\red$ are called reduction relation.}

@@

The reduction function is measurable, and the set of values is measurable, therefore the set of $s$ such that $M$ terminates at $s$ within $n$ steps is measurable for any $n$, therefore $\{s \mid M \text{ terminates at } s \}$ is measurable. 
We say that a term $M$ is \emph{almost-surely terminating} (AST) just if $\mu(\{s \mid M \text{ terminates at } s\}) = 1$.

\iffalse
\lo{Alternatively, define the \emph{runtime of $M$} to be the random variable 
\[
T_M(s) := 
\begin{cases}
\min \set{n \mid \pi_0(\red^n(M, s)) \textrm{ is a value}} & \hbox{if $\red^\infty(M,s)$ is defined}\\
\infty & \hbox{otherwise}
\end{cases}
\]
Equivalently, we say that $M$ is \emph{almost-surely terminating} (AST) if $T_M < \infty$ a.s.; 
and $M$ is \emph{positively almost-surely terminating} (PAST) if $\expect{T_M} < \infty$.}
\fi
For example, the term 
\[
(\tY \lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)} ) \, \underline{0},
\] 
which generates a geometric distribution, terminates on the set $S \setminus [0.5,1]^\mathbb N$, which has measure 1, therefore it terminates almost surely, whereas 
\[
\tif{\tsample - 0.5 < 0}{\underline 0}{(\tY \lambda x. x) \underline 0}
\] 
which terminates on the set $\pi_h^{-1}[[0,0.5]]$, has probability 0.5 of failing to terminate.

>>>>

The language SPCF is a simply-typed lambda calculus with sampling of real numbers from the closed interval $[0,1]$ and unbounded scoring, following \cite{MakOPW21}. 
Types and terms are defined as follows, where $r$ is a real number, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function, and $\Gamma$ is an environment:
\begin{align*}
  \text{\em types } A, B &::= \textsf{R}  \mid  A \to B \\
  \text{\em values } V &::= \lambda x.M  \mid  \underline{r} \\
  \text{\em terms } M, N &::= V  \mid  x  \mid  M_1 \, M_2  \mid  \underline{f}(M_1,\dots ,M_n)  \mid  \tY M \\
   & \qquad \mid  \tif{M < 0}{N_1}{N_2}  \mid  \tsample  \mid  \tscore(M)
\end{align*}
The typing rules are presented in \Cref{fig:typing rules}.
Terms are identified up to $\alpha$-equivalence, as usual. 
The set of all terms is denoted $\Lambda$, and the set of closed terms is denoted $\Lambda^0$.

@@

\begin{align*}
  E[(\lambda x.M) V] & \to E[M[V/x]] \\
  E[\underline f (\underline r_1, \dots , \underline r_n)] & \to E[\underline{f(r_1,\dots,r_n)}] \\
  E[\tY \lambda x. M] & \to E[\lambda z. M[(\tY \lambda x. M)/x] z] \\
  & \qquad \qquad \text{ where $z$ is not free in $s$}\\
  E[\tif{\underline r < 0}{M_1}{M_2}] & \to E[M_1] \text{ where }r < 0 \\
  E[\tif{\underline r < 0}{M_1}{M_2}] & \to E[M_2] \text{ where }r \geq 0 \\
  E[\tsample] & \to E[\underline r] \text{ where } r \in [0,1] \\
  E[\tscore(\underline r)] & \to E[\underline r].
\end{align*}
We write $\to^\ast$ for the reflexive, transitive closure of $\to$.
\lo{The $\tscore$ rule should have a side condition: if $r > 0$.}

@@

\subsection{Sampling semantics}
\label{sec:sampling semantics}
This version of the reduction relation allows $\tsample$ to reduce to any number in $[0,1]$. To more precisely specify the probabilities, an additional argument is needed to determine the outcome of random samples. Let $ I = [0,1] \subset \mathbb{R} $, and let $\entrosp = I^{\mathbb{N}}$, with the \changed[lo]{Borel $\sigma$-algebra $\calF$} and the probability measure, denoted $\mu_{\entrosp}$, given by the limit of $1 \gets I \gets I^2 \gets \cdots$, where the maps are the projections that ignore the last element. Equivalently, a basis of measurable sets is $\prod_{i=0}^\infty X_i$ where $X_i$ are all Borel and all but finitely many are $I$, and $\mu_{\entrosp} (\prod_{i=0}^\infty X_i) = \prod_{i=0}^\infty \Leb(X_i)$.
The maps $\pi_h:\entrosp \to I, \; \pi_t:\entrosp \to \entrosp$ popping the first element are then measurable.
\changed[lo]{Following \cite{DBLP:conf/esop/CulpepperC17}, we call the probability space $(\entrosp, \calF, \mu_{\entrosp})$ the \emph{entropy space}.}

@@

The one-step reduction is given by the function $\red : \Lambda^0 \times \entrosp \to \Lambda^0 \times \entrosp$ where
\begin{equation*}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N\\
        & \qquad \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ is a value}
    \end{array} \right .
\end{equation*}
\lo{Both $\to$ and $\red$ are called reduction relation.}

@@

The reduction function is measurable, and the set of values is measurable, therefore the set of $s$ such that $M$ terminates at $s$ within $n$ steps is measurable for any $n$, therefore $\{s \mid M \text{ terminates at } s \}$ is measurable. 
We say that a term $M$ is \emph{almost-surely terminating} (AST) just if $\mu_{\entrosp}(\{s \mid M \text{ terminates at } s\}) = 1$.

\iffalse
\lo{Alternatively, define the \emph{runtime of $M$} to be the random variable 
\[
T_M(s) := 
\begin{cases}
\min \set{n \mid \pi_0(\red^n(M, s)) \textrm{ is a value}} & \hbox{if $\red^\infty(M,s)$ is defined}\\
\infty & \hbox{otherwise}
\end{cases}
\]
Equivalently, we say that $M$ is \emph{almost-surely terminating} (AST) if $T_M < \infty$ a.s.; 
and $M$ is \emph{positively almost-surely terminating} (PAST) if $\expect{T_M} < \infty$.}
\fi
For example, the term 
\[
(\tY \lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)} ) \, \underline{0},
\] 
which generates a geometric distribution, terminates on the set $\entrosp \setminus [0.5,1]^\mathbb N$, which has measure 1, therefore it terminates almost surely, whereas 
\[
\tif{\tsample - 0.5 < 0}{\underline 0}{(\tY \lambda x. x) \underline 0}
\] 
which terminates on the set $\pi_h^{-1}[[0,0.5]]$, has probability 0.5 of failing to terminate.

### 2020-12-28 LICS21/supermartingales.tex
(Global change: probability space (S, \calF, \mu) -> (\entrosp, \calF, \mu_\entrosp)

\changed[lo]{Fix a probability space $(\Omega, \calF, \mathbb{P})$, and a filtration $(\calF_n)_{n \geq 0}$. 
(In subsequent applications to termination analysis, $\Omega = S$, and $\mathbb{P} = \mu$.)
Let $T$ be a random variable that takes values in $\mathbb{N} \cup \set{\infty}$.
We call $T$ a \emph{stopping time adapted to $(\calF_n)_{n \geq 0}$} just if $\set{T = n} \in \calF_n$, for all $n \geq 0$.}

\begin{definition}\rm
\begin{enumerate}
\item A sequence of random variables $(Y_n)_{n \geq 0}$ adapted to a filtration $(\calF_n)_{n \geq 0}$ is a \emph{supermartingale} if for all $n \geq 0$, $Y_n$ is integrable (i.e.~$\expect{|Y_n|} < \infty$), and $\expect{Y_{n+1} \mid \calF_n} \leq Y_n$ a.s.~\changed[lo]{(i.e.~for all $A \in \calF_n$, $\int_A \mathbb{P}(\dif \omega) \, Y_{n+1}(\omega) \leq \int_A \mathbb{P}(\dif \omega) \,Y_n(\omega)$).}
\item Let $\epsilon > 0$. A stopping time $T$ and a supermartingale $(Y_n)_{n \geq 0}$, both adapted to filtration $(\calF_n)_{n \geq 0}$, are a \emph{$\epsilon$-ranking supermartingale} if for all $n$, $Y_n \geq 0$ and $\expect{Y_{n+1} \mid \calF_n} \leq Y_n - \epsilon \cdot \mathbf{1}_{\set{T > n}}$.\footnote{For $A \in \calF$, we write ${\bf 1}_A$ for the \emph{indicator function} of $A$, i.e., the random variable defined by ${\bf 1}_A(s) := 1$ if $s \in A$, and $0$ otherwise.}
A \emph{ranking supermartingale} is a $\epsilon$-ranking supermartingale for some $\epsilon > 0$.
\end{enumerate}
\end{definition}

@@

\begin{lemma}%[{\citep[Lemma 5.5]{DBLP:conf/popl/FioritiH15}}]
% \citep[Prop 1]{DBLP:conf/popl/ChatterjeeFNH16}}
\label{lem:rank-PAST}
Let $(Y_n)_{n \geq 0}$ be a $\epsilon$-ranking supermartingale with respect to the stopping time $T$, then $T < \infty$ almost surely, and $\expect{T} \leq \frac{\expect{Y_0}}{\epsilon}$.
\end{lemma}

@@

\begin{remark}
\changed[lo]{\Cref{lem:rank-PAST} is a slight extension of \citep[Lemma 5.5]{DBLP:conf/popl/FioritiH15},
which asserts the same result for the specific stopping time $T : s \mapsto \min\set{n \mid Y_n(s) = 0}$.}
\end{remark}

@@

The following is an iterated version of Doob's well-known Optional Sampling Theorem (see, e.g., \cite[\S 6.7]{AshDD00} and \cite[Thm.~7.2]{DBLP:conf/popl/FioritiH15}).
\begin{theorem}[Optional Sampling]
\label{thm:optional sampling}
Let $(X_n)_{n \geq 0}$ be a supermartingale, and $(T_n)_{n \geq 0}$ a sequence of increasing stopping times, both adapted to filtration $(\calF_n)_{n \geq 0}$, then $(X_{T_n})_{n \geq 0}$ is a supermartingale adapted to $(\calF_{T_n})_{n \geq 0}$ if one of the following conditions holds:
\begin{enumerate}
\item each $T_n$ is bounded i.e.~$T_n < c_n$ where $c_n$ is a constant
\item $(X_n)_{n \geq 0}$ is uniformly integrable.
\end{enumerate}
\end{theorem}

@@

%\emph{Notation.}
\changed[lo]{For the rest of this section, fix the probability space $(S, \calF, \mu)$, and a closed SPCF term $M$.
For $n \geq 0$, define the random variables $M_n(s) := \pi_0 (\red^n(M, s))$, and the filtration $(\mathcal{F}_n)_{n \geq 0}$ where $\mathcal{F}_n := \sigma(M_1, \cdots, M_n)$.}

Our first result is the following theorem.
\begin{theorem}[Rankability] 
\label{thm:rankable and strict rankable}
If a closed SPCF term $M$ is rankable (respectively, strictly rankable) by $f$ satisfying $f(N) = 0$ iff $N$ is a value, then $(f(M_n))_{n \geq 0}$ is a supermartingale (respectively, ranking supermartingale with stopping time $T_M: s \mapsto \min \set{n \mid f(M_n(s)) = 0}$) adapted to the filtration $(\mathcal{F}_n)_{n \geq 0}$. %where $\mathcal{F}_n = \sigma(M_1, \cdots, M_n)$.
\end{theorem}

@@

Given an $\omega$-sequence (e.g.~$s \in S$) and $m \geq 0$, we write $s_{\leq m} \in I^m$ to mean the prefix of $s$ of length $m$.
%Fix a closed SPCF term $M \in \Lambda^0$, and a ranking function $f$ for it, satisfying $f(N) = 0$ iff $N$ is a value.
%Let $n \geq 0$, define the following random variables on the probability space $(S, \calF, \mu)$:
%\begin{align*}
%M_n(s) &:= \pi_0 (\red^n(M, s))\\
%\ndraw{n}{s} &:= |\{k < n \mid \exists E: M_k(s) = E[\tsample]\}|,
%\end{align*}
For $n \geq 0$, define
\begin{align*}
%M_n(s) &:= \pi_0 (\red^n(M, s))\\
\ndraw{n}{s} &:= |\{k < n \mid \exists E: M_k(s) = E[\tsample]\}|,
\end{align*}
so that $\pi_1 (\red^n(M, s)) = \overbrace{\pi_t( \cdots (\pi_t}^{\ndraw n s}(s))$. 
%and the filtration $\mathcal{F}_n := \sigma(M_1, \cdots, M_n)$.
The $\calF_n$-measurability of $M_n$ (and hence of $\#\mathrm{draw}_n$) follows from \citep{DBLP:conf/icfp/BorgstromLGS16}.
%$\#\mathrm{draw}_n$ is a stopping time (bounded by $n$). 
\iffalse
\akr{$\#\mathrm{draw}_n$ is not a stopping time. It doesn't look like you actually use this claim anyway, so it should just be fine to remove, but did you mean something different?} \lo{I agree, and I don't actually use this claim.}
\fi
Take $s \in A \in \calF_n$ with $\ndraw{n}{s} = l$.
For any $s'\in S$, if $s_{\leq l} = s_{\leq l}'$ then $s' \in A$.
It follows that ${\set{s_{\leq l}} \cdot I^\omega} \subseteq A$.
\lo{NOTE. To fix notation:
\(
\sigma(M_{n}) = \sigma\big(\set{M_{n}^{-1}(\alpha_j, U_{\alpha_j})
\mid \alpha_j\in \mathsf{Sk}_j, U_{\alpha_j} \in \mathcal{B}(\Real^j), j \geq 0}\big)
\)
}
%Take $M \in \Lambda^0$ (closed SPCF terms). Define, for each $n \in \omega$, the random variable $M_n : S \to \Lambda^0$ by $M_n(s) := \pi_0 (\red^n(M, s))$.

@@

In the rest of this section we show that if $M \in \Lambda^0$ is rankable by $f$, then $M$ is AST.

>>>>

\changed[lo]{Fix a probability space $(\Omega, \calF, \mathbb{P})$, and a filtration $(\calF_n)_{n \geq 0}$. 
(In subsequent applications to termination analysis, $\Omega = S$, and $\mathbb{P} = \mu_{\entrosp}$.)
Let $T$ be a random variable that takes values in $\mathbb{N} \cup \set{\infty}$.
We call $T$ a \emph{stopping time adapted to $(\calF_n)_{n \geq 0}$} just if $\set{T = n} \in \calF_n$, for all $n \geq 0$.}

\begin{definition}\rm
\begin{enumerate}
\item A sequence of random variables $(Y_n)_{n \geq 0}$ adapted to a filtration $(\calF_n)_{n \geq 0}$ is a \emph{supermartingale} if for all $n \geq 0$, $Y_n$ is integrable (i.e.~$\expect{|Y_n|} < \infty$), and $\expect{Y_{n+1} \mid \calF_n} \leq Y_n$ a.s.~\changed[lo]{(i.e.~for all $A \in \calF_n$, $\int_A \mathbb{P}(\dif \omega) \, Y_{n+1}(\omega) \leq \int_A \mathbb{P}(\dif \omega) \,Y_n(\omega)$).}
\item Let $\epsilon > 0$. 
Given a stopping time $T$ and a supermartingale $(Y_n)_{n \geq 0}$, both adapted to filtration $(\calF_n)_{n \geq 0}$, 
we say that $(Y_n)_{n \geq 0}$ is a \emph{$\epsilon$-ranking supermartingale w.r.t.~$T$} if for all $n$, $Y_n \geq 0$ and $\expect{Y_{n+1} \mid \calF_n} \leq Y_n - \epsilon \cdot \mathbf{1}_{\set{T > n}}$.
{(For $A \in \calF$, we write ${\bf 1}_A$ for the \emph{indicator function} of $A$, i.e., the random variable defined by ${\bf 1}_A(\omega) := 1$ if $\omega \in A$, and $0$ otherwise.)} \lo{\LaTeX\ does not allow the preceding bracketed sentence to be typeset as a footnote. Strange!}
A \emph{ranking supermartingale} w.r.t.~$T$ is a $\epsilon$-ranking supermartingale w.r.t.~$T$ for some $\epsilon > 0$.
\end{enumerate}
\end{definition}

@@

\begin{lemma}%[{\citep[Lemma 5.5]{DBLP:conf/popl/FioritiH15}}]
% \citep[Prop 1]{DBLP:conf/popl/ChatterjeeFNH16}}
\label{lem:rank-PAST}
Let $(Y_n)_{n \geq 0}$ be a $\epsilon$-ranking supermartingale w.r.t.~the stopping time $T$, then $T < \infty$ a.s., and $\expect{T} \leq \frac{\expect{Y_0}}{\epsilon}$.
\end{lemma}

@@

\begin{remark}
\changed[lo]{\Cref{lem:rank-PAST} is a slight extension of \citep[Lemma 5.5]{DBLP:conf/popl/FioritiH15},
which asserts the same result but for the specific stopping time $T : \omega \mapsto \min\set{n \mid Y_n(\omega) = 0}$.}
\end{remark}

@@

The following is an iterated version of Doob's well-known Optional Sampling Theorem (see, e.g., \cite[\S 6.7]{AshDD00} and \cite[Thm.~7.2]{DBLP:conf/popl/FioritiH15}).
\begin{theorem}[Optional Sampling]
\label{thm:optional sampling}
Let $(X_n)_{n \geq 0}$ be a supermartingale, and $(T_n)_{n \geq 0}$ a sequence of increasing stopping times, all adapted to filtration $(\calF_n)_{n \geq 0}$, then $(X_{T_n})_{n \geq 0}$ is a supermartingale adapted to $(\calF_{T_n})_{n \geq 0}$ if one of the following conditions holds:
\begin{enumerate}
\item each $T_n$ is bounded i.e.~$T_n < c_n$ where $c_n$ is a constant
\item $(X_n)_{n \geq 0}$ is uniformly integrable.
\end{enumerate}
\end{theorem}

@@

%\emph{Notation.}
\changed[lo]{Henceforth, fix the probability space $(S, \calF, \mu_{\entrosp})$, and a closed SPCF term $M$.
For $n \geq 0$, define the random variables $M_n(s) := \pi_0 (\red^n(M, s))$, and the filtration $(\mathcal{F}_n)_{n \geq 0}$ where $\mathcal{F}_n := \sigma(M_1, \cdots, M_n)$.}

Our first result is the following theorem.
\begin{theorem}[Rankability] 
\label{thm:rankable and strict rankable}
If a closed SPCF term $M$ is rankable (respectively, strictly rankable) by $f$ satisfying $f(N) = 0$ iff $N$ is a value, then $(f(M_n))_{n \geq 0}$ is a supermartingale (respectively, ranking supermartingale w.r.t.~stopping time $T_M: s \mapsto \min \set{n \mid f(M_n(s)) = 0}$) adapted to $(\mathcal{F}_n)_{n \geq 0}$. %where $\mathcal{F}_n = \sigma(M_1, \cdots, M_n)$.
\end{theorem}

@@

Given an $\omega$-sequence (e.g.~$s \in S$) and $m \geq 0$, we write $s_{\leq m} \in I^m$ to mean the prefix of $s$ of length $m$.
%Fix a closed SPCF term $M \in \Lambda^0$, and a ranking function $f$ for it, satisfying $f(N) = 0$ iff $N$ is a value.
%Let $n \geq 0$, define the following random variables on the probability space $(S, \calF, \mu)$:
%\begin{align*}
%M_n(s) &:= \pi_0 (\red^n(M, s))\\
%\ndraw{n}{s} &:= |\{k < n \mid \exists E: M_k(s) = E[\tsample]\}|,
%\end{align*}
For $n \geq 0$, define
\begin{align*}
%M_n(s) &:= \pi_0 (\red^n(M, s))\\
\ndraw{n}{s} &:= |\{k < n \mid \exists E: M_k(s) = E[\tsample]\}|
\end{align*}
so that $\pi_1 (\red^n(M, s)) = %\overbrace{\pi_t( \cdots (\pi_t}^{\ndraw n s}(s))$. 
\pi_t( \cdots (\pi_t (s))$, with $\pi_t$ applied ${\ndraw n s}$ times.
%and the filtration $\mathcal{F}_n := \sigma(M_1, \cdots, M_n)$.
The $\calF_n$-measurability of $M_n$ (and hence of $\#\mathrm{draw}_n$) follows from \citep{DBLP:conf/icfp/BorgstromLGS16}.
%$\#\mathrm{draw}_n$ is a stopping time (bounded by $n$). 
\iffalse
\akr{$\#\mathrm{draw}_n$ is not a stopping time. It doesn't look like you actually use this claim anyway, so it should just be fine to remove, but did you mean something different?} \lo{I agree, and I don't actually use this claim.}
\fi
Take $s \in A \in \calF_n$ with $\ndraw{n}{s} = l$.
For any $s'\in S$, if $s_{\leq l} = s_{\leq l}'$ then $s' \in A$.
It follows that ${\set{s_{\leq l}} \cdot I^\omega} \subseteq A$.
\lo{NOTE. To fix notation:
\(
\sigma(M_{n}) = \sigma\big(\set{M_{n}^{-1}(\alpha_j, U_{\alpha_j})
\mid \alpha_j\in \mathsf{Sk}_j, U_{\alpha_j} \in \mathcal{B}(\Real^j), j \geq 0}\big)
\)
}
%Take $M \in \Lambda^0$ (closed SPCF terms). Define, for each $n \in \omega$, the random variable $M_n : S \to \Lambda^0$ by $M_n(s) := \pi_0 (\red^n(M, s))$.

@@

In the rest of this section we show that if $M \in \Lambda^0$ is rankable, then $M$ is AST: in other words, the method of ranking function is sound for proving a.s.~termination of SPCF programs.
